---
title: "Stator-analysis"
output: html_document
date: "2024-03-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the packages

```{r library,echo=FALSE}
library(gridExtra, verbose=FALSE)        
library(RColorBrewer, verbose=FALSE)
library(ComplexHeatmap)
library(biomaRt)
library(Seurat)
library(stringr)
library(pheatmap)
library(dplyr)
library(rrvgo)
library(DT)
library(clusterProfiler)
library(data.table)
library(ggplot2)
library(igraph)
library(ggrepel)
library(org.Hs.eg.db)
library(org.Mm.eg.db)
library("limma")
library(heatmaply)
library(plotly)
```

# Table-Tab: Filter Dtuples and clustering

Create D-tuples clustering results with the following codes:

```{r Produce the result files, echo=FALSE}
cutoff="Optimal" # or a number from 0-1
data_path_devStates="./data/all_DTuples.csv"
data_path_trainData="./data/trainingData_14698Cells_1000Genes.csv"
minStateDeviation=3
minNoCells=0
stateDevAlpha=0.05
command=paste("python ./Produce_devStates.py",cutoff,data_path_devStates,data_path_trainData,minStateDeviation,minNoCells,stateDevAlpha,sep = " ")
print(command)
print("Type the above command line will produce:")
print(paste0(cutoff,"_",minStateDeviation,"_",minNoCells,"_",stateDevAlpha,'_devStates.csv'))
print(paste0(cutoff,"_",minStateDeviation,"_",minNoCells,"_",stateDevAlpha,'_binReps.csv'))
print(paste0(cutoff,"_",minStateDeviation,"_",minNoCells,"_",stateDevAlpha,'_modularity_scores.csv'))
```

Read in the created file:

```{r Show the dtuples clustering table}
Devstates<-read.csv(paste0('./',cutoff,"_",minStateDeviation,"_",minNoCells,"_",stateDevAlpha,'_devStates.csv'),colClasses = c("character"))
Devstates$enrichment<-as.numeric(Devstates$enrichment)
Devstates$cluster<-as.numeric(Devstates$cluster)
Devstates$pval_corrected<-as.numeric(Devstates$pval_corrected)
Devstates$No.Cells<-as.numeric(Devstates$No.Cells)
Devstates<-Devstates[order(Devstates$cluster),]
Devstates<-Devstates[,-1]
head(Devstates)
dim(Devstates)
table(Devstates$cluster)
```

Get the cells in each stator state:

```{r Mapping the cells with count matrix}
source("./app/tabs/heatmap/heatmap.R") # use the getGetCellList function
load("./data/Example.Rdata") # load the binarised count matrix

# count<-Count_matrix
# count[count>0]<-1
# count[count==0]<-0

count[1:10,1:10]
head(Devstates)
List<-GetCellList(count,Devstates)

head(List,2)
lengths(List)

List_dtuples<-GetCellList_d(count,Devstates)
head(List,2)
lengths(List)
table(lengths(List_dtuples)==Devstates$No.Cells)
```

# Heatmap Cells: Perform over/under-representation test

```{r Over/under representation test with meta data information}

# Heatmap for cells(Tab)
Meta_data_path<-"./data/Meta_Data.csv"
Meta_data<-read.csv(Meta_data_path,row.names = 1,colClasses = "character")

TestSelected_Cells<-"log10FDR" 
# Over_representation/Under_representation: log10FDR
# or Fold for Fisher test

colorCells<-colorRampPalette(c("white","firebrick3"))(100) # Over_representation
# Under_representation:colorRampPalette(c("white","#2166ac"))(100)}
# Two-side Fisher: colorRampPalette(c("#2166ac", "white", "#cc3333"))(100)

colnames(Meta_data)<-c("Cell_State","Cell_Types") 
# Cell_Types refer to the second column here 
# heatmap() with test for the second column and stator state

# Test results
Test="Over_representation" 
#or Under_representation or Fisher

# use the heatmap function
result1<-heatmap(Meta_data,Devstates,List,N=dim(count)[1],Test)
result1$log10FDR[1:10,1:10]

# Add the title
result1$cutoff=cutoff

```

Plot the results from over/under-representation test for cells in each stator state

```{r Plot heatmap for results}
#    

if (TestSelected_Cells=="log10FDR"){
  
p<-ComplexHeatmap::pheatmap(border_color = NA,result1[[TestSelected_Cells]],display_numbers = result1[["Mydata_raw_m"]],
top_annotation = columnAnnotation(Pct=anno_barplot(border = F,result1[["col"]])),  
left_annotation = rowAnnotation(Pct=anno_barplot(border = F,result1[["row"]])),
show_column_dend = FALSE, show_row_dend = FALSE,
fontsize = 12,fontsize_number = 15, name ="-Log10FDR",heatmap_legend_param=list(title_position = "lefttop-rot"),
main=result1[["cutoff"]],breaks = seq(0, 10, by = 0.1),
cluster_cols = T,cluster_rows = T,treeheight_row = 0, treeheight_col = 0,color = colorCells)

} else{    # Fold
  
p<- ComplexHeatmap::pheatmap(border_color = NA,result1[[TestSelected_Cells]],display_numbers = result1[["Mydata_raw_m"]],
 top_annotation = columnAnnotation(Pct=anno_barplot(border = F,result1[["col"]])),  
left_annotation = rowAnnotation(Pct=anno_barplot(border = F,result1[["row"]])),
show_column_dend = FALSE, show_row_dend = FALSE,     
fontsize = 12,fontsize_number = 15,name = "Log10 Odds ratio",heatmap_legend_param=list(title_position = "lefttop-rot"),
  main=result1[["cutoff"]],
cluster_cols = T,cluster_rows = T,treeheight_row = 0, treeheight_col = 0,color = colorCells,breaks = seq(-2, 2, length.out = 100))

}
      
draw(p,heatmap_legend_side = "left")

```

Plot interactive heatmap-cells

```{r interactive heatmap}
# Heatmaply
mt1<-result1[["Overlap"]]
mt2<-result1[["Mydata_raw_m"]]
mt2[mt2=="*"]<-"Yes"
mt2[mt2==" "]<-"No"
matr<-array(data = NA,dim=dim(mt1))
for (i in 1:nrow(matr)){
        matr[i,]<-paste0("Overlap is: ",mt1[i,],"\nSignificant (FDR<0.05): ",mt2[i,])
}

if (!TestSelected_Cells=="Fold"){
heatmaply(
          result1[[TestSelected_Cells]], colors = colorCells,show_dendrogram=c(F,F),
          Rowv=pheatmap(result1[[TestSelected_Cells]])[[1]],Colv=rev(pheatmap(result1[[TestSelected_Cells]])[[2]]),
          custom_hovertext=matr,method = "ggplot")
        } else{
        heatmaply(
          result1[[TestSelected_Cells]],colors = colorCells,scale_fill_gradient_fun = ggplot2::scale_fill_gradient2(low = "#2166ac",high = "#cc3333",midpoint = 0,mid = "white"),Colv=rev(pheatmap(result1[[TestSelected_Cells]])[[2]]),Rowv=pheatmap(result1[[TestSelected_Cells]])[[1]],
          custom_hovertext=matr,show_dendrogram=c(F,F))
      }

```

# Heatmap Genes: Perform over/under-representation test for d-tuple genes in each stator state

```{r heatmap-cells}
source("./app/tabs/GO_plot/GO_plot.R")
source("./app/tabs/heatmap_gene/heatmap_gene.R")
usedCellStateGene<-"./data/CancerState.csv"

TestSelected_Genes<-"log10FDR" #or Fold 

Test_Gene="Over_representation" # or Under_representation or Fisher

GetGenesList<-GetGenes(Devstates,Remove=T) # remove 
GetGenesList_All<-GetGenes(Devstates,Remove=F) # use all genes
    

colorGenes<-colorRampPalette(c("white","firebrick3"))(100)
# colorRampPalette(c("white","#2166ac"))(100) for under-representation test
# colorRampPalette(c("#2166ac", "white", "#cc3333"))(100) for fisher two side test
# N the number of background
result_genes<-heatmapGenes(usedCellStateGene,GetGenesList_All,N=25678,Test_Gene)
result_genes$cutoff=cutoff
```

Plot heatmap

```{r plot heatmap gene}
#Plot heatmap
if (!TestSelected_Genes=="Fold"){

p<-ComplexHeatmap::pheatmap(border_color = NA,result_genes[[TestSelected_Genes]],display_numbers = result_genes[["Mydata_raw_m"]],fontsize = 12,fontsize_number = 15,name = "-Log10FDR",heatmap_legend_param=list(title_position = "lefttop-rot"),top_annotation = columnAnnotation(Pct=anno_barplot(border = F,result_genes[["col"]])),  
left_annotation = rowAnnotation(Pct=anno_barplot(border = F,result_genes[["row"]])),
show_column_dend = FALSE, show_row_dend = FALSE, 
main=result_genes[["cutoff"]],
cluster_cols = T,cluster_rows = T,treeheight_row = 0, treeheight_col = 0,
color =  colorGenes,breaks = seq(0, 10, by = 0.1))

} else{
        
p<-ComplexHeatmap::pheatmap(border_color = NA,result_genes[[TestSelected_Genes]],display_numbers = result_genes[["Mydata_raw_m"]],
                                 top_annotation = columnAnnotation(Pct=anno_barplot(border = F,result_genes[["col"]])),  
                                 left_annotation = rowAnnotation(Pct=anno_barplot(border = F,result_genes[["row"]])),
                                 show_column_dend = FALSE, show_row_dend = FALSE,  
                                fontsize = 12,fontsize_number = 15, name = "Log10 Odds ratio",heatmap_legend_param=list(title_position = "lefttop-rot"),
                                main=result_genes[["cutoff"]],
                                cluster_cols = T,cluster_rows = T,treeheight_row = 0, treeheight_col = 0,
                                 color =  colorGenes,  breaks = seq(-2, 2, length.out = 100))

}
      #breaks = seq(0, 10, by = 1)
      draw(p,heatmap_legend_side = "left")
      
```

Plot interactive heatmap-genes

```{r interactive heatmap}

## heatmaply
mt1<-result_genes[["Overlap"]]
mt2<-result_genes[["Mydata_raw_m"]]
mt2[mt2=="*"]<-"Yes"
mt2[mt2==" "]<-"No"
matr<-array(data = NA,dim=dim(mt1))
for (i in 1:nrow(matr)){
        matr[i,]<-paste0("Overlap is: ",mt1[i,],"\nSignificant (FDR<0.05): ",mt2[i,])
      }
      

if (!TestSelected_Cells=="Fold"){
        heatmaply(
          result_genes[[TestSelected_Genes]], colors = colorGenes,show_dendrogram=c(F,F),
          Rowv=pheatmap(result_genes[[TestSelected_Genes]])[[1]],Colv=rev(pheatmap(result_genes[[TestSelected_Genes]])[[2]]),
          custom_hovertext=matr,method = "ggplot")
  
} else{
        heatmaply(
result_genes[[TestSelected_Genes]],colors = colorGenes,scale_fill_gradient_fun = ggplot2::scale_fill_gradient2(low = "#2166ac",high = "#cc3333",midpoint = 0,mid = "white"),Colv=rev(pheatmap(result_genes[[TestSelected_Genes]])[[2]]),Rowv=pheatmap(result_genes[[TestSelected_Genes]])[[1]],
          custom_hovertext=matr,show_dendrogram=c(F,F)) }
```

# GO&KEGG/Using rrvgo for d-tuple genes of each stator states

```{r GO and KEGG}
source("./app/tabs/rrvgo/rrvgo.R")
selected_cluster="1,3" #seperate by number
Mart="hsapiens_gene_ensembl"
go_species="org.Hs.eg.db"
kegg_species="hsa"
BackgroundGenes=colnames(count)[colSums(count)>0]

# Gene enrichment analysis for dtuple genes
GO_KEGG <- FunctionE(selected_cluster,Mart,kegg_species,go_species,GetGenesList,BackgroundGenes)
table(GO_KEGG$class)

p<-Plot_enrichment(GO_KEGG)
p


# rrvGO 
selected_cluster="5"
subOntology="BP" #CC or MF
rrvGO<-rrvgo_FunctionE(selected_cluster,subOntology,go_species,Mart,GetGenesList,BackgroundGenes)

wordcloudPlot(rrvGO$reducedTerms)
treemapPlot(rrvGO$reducedTerms)
scatterPlot(labelSize = 5,rrvGO$simMatrix, rrvGO$reducedTerms)
```

# Upset Plot

```{r upset plot}
source("./app/tabs/upsetplot/upsetplot.R")
selected_cluster_upset="19,46,47"
UpsetMode="intersect" #"distinct", "union"
m<-Up_set(selected_cluster_upset,List,UpsetMode)

ComplexHeatmap::UpSet(m,top_annotation = upset_top_annotation(m, add_numbers = TRUE),column_title =cutoff,right_annotation = upset_right_annotation(m, add_numbers = TRUE),comb_order = order(comb_size(m)))
```

# DE analysis

```{r }
source("./app/tabs/DE/DE.R")
srt # Seurat object 
# can be process by the following:
# Count_matrix<-t(Count_matrix) 
# preprocessing srt object
# print("Preprocessing srt object")
# srt<-CreateSeuratObject(counts =Count_matrix)
# srt <- NormalizeData(srt, normalization.method = "LogNormalize", scale.factor = 10000)
# srt <- FindVariableFeatures(srt, selection.method = "vst", nfeatures = 2000)
# srt <- ScaleData(srt)

# Select cells in group1
selected_clusterDE1<-"46,19"

# Select cells in group2
selected_clusterDE2<-"45,28"

logfc=0.25
Pvalue_DE=0.05

DE_pairwise<-DE_set(selected_clusterDE1,selected_clusterDE2,srt,logfc,Pvalue_DE,List)

p_heatmap<-PlotDEheatmap(selected_clusterDE1,selected_clusterDE2,DE_pairwise,srt,List)

doheatmap<-DoHeatmap(p_heatmap[["sub"]],features = p_heatmap[["top20"]],slot = "c",group.by = "state",
                     disp.min = 0,disp.max = 4) +scale_fill_gradient2( low = rev(c('#d1e5f0','#67a9cf','#2166ac')), mid = "white", high = rev(c('#b2182b','#ef8a62','#fddbc7')), midpoint = 0, guide = "colourbar", aesthetics = "fill")
        
print(doheatmap)

# Enrichment anlaysis for DEGs
DEKEGGO<-DEGO(DE_pairwise,Mart,kegg_species,go_species,logfc,Pvalue_DE, BackgroundGenes)

# Plot
Plot_DE_enrichment(DEKEGGO)
```

# Find Marker Tab

```{r markers}
source("./app/tabs/marker/marker.R")
selected_clusterMarker="6"
logfcMarker=0.25
Pvalue_Marker=0.05

# Identify markers
Marker <- Marker_set(selected_clusterMarker,srt,logfcMarker,Pvalue_Marker,List)

# PlotVolcano  
PlotVolcano(selected_clusterMarker,Marker,List,GetGenesList_All)

# Enrichment analysis
Marker_enrichment<-MarkerGO(Marker,Mart,kegg_species,go_species,logfcMarker,Pvalue_Marker, BackgroundGenes)

# Plot
Plot_Marker_enrichment(Marker_enrichment)

```

# Automatic annotations

```{r Automatic annotations}
source("./app/tabs/automatic/automatic.R")
Annotations_path<-"./data/CancerState.csv"
Annotations<-read.csv(Annotations_path)
regulation="Up_regulated genes"


# Markers for all states
AllMarkers<-MarkerAll_set(srt,logfcMarker,Pvalue_Marker,List)

# Markers in AnnotTable
Markers_in_AnnotTable<- Markers_in_Annot(AllMarkers,Annotations, regulation)

# number all
NumberAllAnnot<- number_all(Annotations)

# summary_table
summaryDF<- summary_df(Markers_in_AnnotTable,NumberAllAnnot)
summaryDF_per<- percent_summaryDF(summaryDF,NumberAllAnnot)
   
#Get gene list in each annotation
summary_genelist<-get_genelist_annotation(Markers_in_AnnotTable,NumberAllAnnot)


# Interactive heatmap
heatmaply(summaryDF, colors = colorRampPalette(c("white","firebrick3"))(100),show_dendrogram=c(F,F),
custom_hovertext=summary_genelist,method = "ggplot",margins = c(5,10,30),dendrogram = "none")

```

# Markov Blanket

```{r markove blanket}
source("./app/tabs/markovBlanket/markovBlanket.R")
# Get markov blanket:

NodeGene<-c("IGHG4")
MCMCGraph_path<-"./data/MCMCgraph_14698Cells_1000Genes.csv"
usedMCMCGraph<-read.csv(MCMCGraph_path,header=T,row.names = 1,colClasses = "character")
usedMCMCGraph<-graph_from_adjacency_matrix(as(usedMCMCGraph, 'matrix'), mode="directed")
      
Markov_blanket<-mb(usedMCMCGraph,NodeGene)

      
subgraph<-subgraph(usedMCMCGraph, unique(c(unlist(Markov_blanket),NodeGene)))
V(subgraph)$color<-"Navy"
V(subgraph)$color[V(subgraph)$name%in%NodeGene]<-"red"

plot(subgraph,layout=layout_with_kk(subgraph),edge.arrow.size = 0.3,vertex.label.color=V(subgraph)$color,vertex.shape="none",vertex.label.font=0.4)
      
```

# UMAP

```{r umap}
source("./app/tabs/umap/umap.R")
selected_umap="4"
UMAP_coords_path<-"./data/UMAP_coords.csv"
usedUMAP_coords<-read.csv(UMAP_coords_path,header=T,row.names = 1)
cluster_umap(selected_umap,usedUMAP_coords,srt,List,Devstates)

```


```{r dendrogram}
library(dendextend)
source("./app/tabs/dendrogram/dendrogram.R")

data_path2="Optimal_3_0_0.05_binReps.csv"

# The optimal dice distance is 0.85 from Optimal_3_0_0.05_modularity_scores.csv
cutoff=0.85
#"read Optimal_3_0_0.05_binReps.csv"

binReps_mat<-binReps(cutoff,data_path2,minStateDeviation,minNoCells,stateDevAlpha)

# Construct tree
hcl<-hcl_construct(binReps_mat,Devstates)

# Get clusters
hcl_cluster<-hcl_cluster(hcl,cutoff,Devstates)

# Color the dend
avg_col_dend<-plot_full(hcl, cutoff,hcl_cluster)

# Plot
par(mar = c(25, 2.5, 0.5, 0.5))
plot(avg_col_dend);rect.hclust(hcl , h = cutoff)

# Get sub dend
selected_dendrogram=45

dend<-as.dendrogram(hcl)
dend_list<-GetSubend(dend,hcl_cluster)
sub<-dend_list[[selected_dendrogram]]

# Plot sub
par(mar = c(15, 5, 1, 1));plot(sub)

```

In the end, it is better to remove the unwanted the figures/files in the directory.
